---
title: "Using quickinspectR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using quickinspectR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

**quickinspectR** is a package designed for beginner R programmers that makes graphically inspecting your data faster and easier. In this article, we'll go through basic usage, intermediate usage, and advance usage of **quickinspectR** using a tried-and-true dataset among R programmers: `palmerpenguins::penguins`. 

# Getting started

To get started, we first need load the **quicksinspectR** package into our session. If you don't already have it installed, you can use the commented out code below to install the `devtools` package - this lets us install the development version of **quickinspectR** straight from GitHub. 

```{r setup}
# install.packages("devtools")
# devtools::install_github("andrewfullerton/quickinspectR")
library(quickinspectR)
```

We also need to load `palmerpenguins` in order to access the `penguins` dataset. Let's do that!

```{r}
# install.packages("palmerpenguins")
library(palmerpenguins)
```

Great! Now that we're set up, we can see how **quickinspectR** works. 

# Basic usage

By design, all **quickinspectR** functions require only one argument: a data frame or a tibble (data). Unless otherwise specified, functions will display all the relevant variables contained in the data frame and default to easy-to-read plot styling.

Currently, **quickinspectR** contains three `inspect` functions:

-   `inspect_normality`: graphically inspect the distribution of numeric variables in your data.
-   `inspect_balance`: graphically inspect class imbalance among categorical variables.
-   `inspect_missing`: graphically inspect missingness in your data.

Each of these functions can be used to help you familiarize yourself with your data and get your analyses off the ground just a bit quicker. 

## Using `inspect_normality` to inspect distribution

Let's say I'm interested in exploring the relationship between body mass, bill length, and flipper length using the `penguins` dataset. One of the first things I likely want to know before I start trying to do any intensive analysis is "are these variables normally distributed?" - the answer to this question will dictate both what types of analysis I can do and how I will approach my analysis.

I can fumble around with `dplyr`, `tidyr`, and `ggplot2` for an hour trying to build the plot I need from the ground-up, or I can `inspect_normality` and get to the fun stuff sooner.

```{r}
inspect_normality(penguins)
```

In a single line of code, I have a pretty good sense of how these variables are distributed. If I want to get rid of some of the clutter in that plot, I can tell `inspect_normality` which variables I'm interested in using the `vars` argument.

```{r}
inspect_normality(data = penguins,
                  vars = c("bill_length_mm", "flipper_length_mm", "body_mass_g"))
```

Using the `vars` argument, I can focus on the variables were most interested in and leave the variables I'm not interested in out of the plot. 

## Using `inspect_balance` to inspect class imbalance

Now that I've inspected some numeric variables, I'm thinking that there might be some connection between the `species` of a penguin, its `sex`, and its physical attributes. There are a few important details I want to know right away. How many different species am I dealing with in this dataset? Are there enough observations corresponding to each classification of species/sex for me to use them meaningfully in my analysis?

Before undertaking any formal analysis, I want to make sure I'm well-informed about this. I can `inspect_balance` to figure this out quickly.

```{r}
inspect_balance(penguins)
```

With a single line of code, I can see that there's a pretty even split between male and female penguins, but there's much less balance between the different species of penguins. I'll keep this in mind before I start trying to build any models.

## Using `inspect_missing` to identify missing values

You may have noticed that an `NA` found its way into the above plot. This is a good reminder that I should take a look to see if there's any missing data in the `penguins` dataset. If data is missing in a patterned way, it could introduce bias into my analysis. I can `inspect_missing` to figure this out.

```{r}
inspect_missing(penguins)
```

Most of the missing data does seem to be in the `sex` variable. I think it'll be okay, but I'll keep this in mind if I start seeing some unexpected results later on. And just like we did before, I can use `vars` to select only those variables I'm most interested in inspecting. 

```{r}
inspect_missing(data = penguins,
                vars = c("bill_length_mm", "flipper_length_mm", "body_mass_g", "species", "sex"))
```

Now that I have my bearings, I can move on to more advanced (and time-intensive) analysis. Using **quickinspectR** I was able to accomplish a few routine inspections in just a few minutes with just a few easy lines of code.

# Intermediate usage

If you wanted to, you could stop reading right now and go use **quickinspectR** in your work, but sometimes it's nice to add a little pop of colour to your visualizations. 

## Plot customization

All `inspect` functions in **quickinspectR** contain several additional (but optional) arguments to make plot customization more accessible for the beginner (or the experienced programmer who's in a rush).

-   `vars` (as seen above) allows you to easily select the variables you want to visualize.
-   `fill_colour` lets you pick your colour of choice.
-   `title` lets you add a descriptive title to your plot.

```{r}
inspect_normality(data = penguins,
                  vars = c("bill_length_mm", "flipper_length_mm", "body_mass_g"),
                  fill_colour = "blue",
                  title = "Penguin Attributes")
```

Now you can inspect your data *in style*! While some `inspect` functions contain additional arguments, all share these same basic customization options. To see what additional arguments a particular `inspect` function has, you can check the documentation by running `?inspect_...` in the console, replacing `...` with your function of choice. For example, running `?inspect_normality` will reveal an additional `bins` argument that lets you specify the number of bins to use in each histogram visualization.

# Advanced usage

While **quickinspectR** is intended as a package to help novice R programmers get their analyses off the ground more efficiently rather than build deeply customized plots, there are plenty of opportunities to extend **quickinspectR** when you want to.

## Using `...` to extend `inspect` functions

Under the hood, **quickinspectR** is generating **ggplot2** objects. Using the `...` argument, you can pass any supported ggplot2 argument directly into the plot's "main" `geom_` layer. When using `inspect_normality`, your arguments will be passed right into the `geom_histogram` layer. When using `inspect_balance` and `inspect_missing`, your arguments will land in the `geom_bar` layer.

```{r}
inspect_normality(data = penguins,
                  vars = c("bill_length_mm", "flipper_length_mm", "body_mass_g"),
                  fill_colour = "navyblue",
                  title = "Penguin Attributes",
                  colour = "lightblue", # Specifies the bar outline colour
                  alpha = 0.7) # Reduces the opacity
```

Using `colour` and `alpha`, we were able to add additional customizations to our plot (beyond those explicitly supported by `inspect_normality`).

## Adding additional ggplot2 layers

While passing additional arguments into a pre-specified `geom` layer is great, it's not without limitations. Since `inspect` functions return a standard ggplot2 object, you can easily add themes, layers, and more using standard `ggplot2` syntax. Let's add a dark theme to the plot.

```{r}
library(ggplot2) # Load ggplot2

inspect_normality(data = penguins,
                  vars = c("bill_length_mm", "flipper_length_mm", "body_mass_g"),
                  fill_colour = "navyblue",
                  title = "Penguin Attributes",
                  colour = "lightblue",
                  alpha = 0.7) + 
  theme_dark() # Adds dark theme
```

Another situation where this comes in handy is when you want to use an `inspect` function to get your plot *most* of the way there, but add some additional modifications and `geom` layers based on a specific use case. Let's add some custom axis labels and a `geom_rug` layer.

```{r}
inspect_normality(data = penguins,
                  vars = c("bill_length_mm", "flipper_length_mm", "body_mass_g"),
                  fill_colour = "navyblue",
                  title = "Penguin Attributes",
                  colour = "lightblue",
                  alpha = 0.7) +
  theme_dark() + 
  labs(x = "Physical attributes", y = "Count") + # Axis labels
  geom_rug(aes(x = value), color = "red", alpha = 0.5) # Geom layer
```

# Happy inspecting!

There you have it! In this article, we've covered some basic use cases, some intermediate use cases, and even a couple advanced use cases that push the limits of **quickinspectR**. Happy inspecting!
